---
title: "Generalized Linear Model (GLM)"
author: "Emma Li"
date: March 27, 2017
output: slidy_presentation
---


# Compare Linear Regression & GLM

Linear Regression: E(Y) = $\mu$ = X * $\beta$ | GLM: g(E(Y)) = g($\mu$) = X * $\beta$
----------------------------------------------|-----------------------------------------------------------------
* Linear relationship between X and E(Y)      |* Linear relationship between X and g( E(Y) )
* Multivariate normality                      |
* No or little multicollinearity              |* No or little multicollinearity
* No auto-correlation                         |* No auto-correlation
* Error terms have similar variances          |* Error terms have similar variances 

# Review GLM Examples

* Distributions in Exponential Family: Normal, Bernoulli, Binomial, Poisson, Negative Binomial, Gamma, Tweedie, Exponential, etc. 

1. Y is count (e.g. claim count): Poisson distribution

2. Y is binary (e.g., loss or no loss): Bernoulli distribution

* Link Functions

1. Poisson distribution: log function = ln($\lambda$) = X*$\beta$

2. Bernoulli distribution: logit function = ln( p/(1-p) ) = X*$\beta$


# R Packages

1. stats: glm() is used to fit generalized linear models

2. insuranceData: 'A Collection of Insurance Datasets Useful in Risk Classification in Non-life Insurance'

* Inputs: glm(formula, family, data, ...)

* Outputs:  coefficients, p values, residuals, fitted values, summary, ...


# Case Study
One dataset in package 'insuranceData' is called 'dataCar' and it is based on one-year vehicle insurance policies taken out in 2004 or 2005. 

Variables     | Descriptions
------------- | -------------
numclaims	    | number of claim
claimcst0     | claim amount 
veh_body      | vehicle type
veh_age       | vehicle age
gender        | driver gender
area          | location
agecat        | driver age

# Summary and Graphs
```{r size = "tiny"}
library("insuranceData")
data(dataCar)
```

```{r }
dim(dataCar)
colnames(dataCar)
```

# Summary and Graphs
```{r }
head(dataCar)
```

# Summary and Graphs
```{r }
str(dataCar)
```

# Summary and Graphs
```{r}
summary(dataCar)
```

# Summary and Graphs
```{r}
table(dataCar$numclaims)
hist(dataCar$numclaims)
```

# Basic Data Manipulation
```{r  }
avg <- function(x) {
  dat <- aggregate(dataCar$numclaims, by = list(dataCar[, x]), FUN = mean)
  barplot(dat$x, xlab = x, ylab = "Claim Count Averages")
  axis(side=1, at=1:nrow(dat), labels=dat$Group.1)
}

par(mfrow=c(2,2))
  avg(x = "veh_age")
  avg(x = "area")
  avg(x = "agecat")
  avg(x = "veh_body")
```

# Basic Data Manipulation
```{r }
summaries<-function(x) {
  means<-aggregate(dataCar$numclaims, by = list(dataCar[, x]), FUN = mean)
  lengths<-aggregate(dataCar$numclaims, by = list(dataCar[, x]), FUN = length)
  means_lengths<-merge(means,lengths,by="Group.1"  )
  colnames(means_lengths)<-c(x,'numclaims average','count')
  means_lengths[order(means_lengths[,"numclaims average"]),]
}
summaries("veh_body")
```

```{r }
library("plyr")
dataCar$veh_body_cat<-mapvalues(dataCar$veh_body
                      , from = c("COUPE", "MCARA", "MIBUS", "PANVN", "RDSTR")
                      , to = c("Others","Others","Others","Others","Others"))
summaries("veh_body_cat")

par(mfrow=c(2,2))
  avg(x = "veh_body")
  avg(x = "veh_body_cat")
```

# Basic Data Manipulation
```{r }
summaries("veh_age")
```

```{r }
dataCar$veh_age_cat<-factor(ifelse(dataCar$veh_age =='2','2','Others'))
aggregate(dataCar$numclaims, by = list(dataCar[, "veh_age_cat"]), FUN = mean)
summaries("veh_age_cat")

par(mfrow=c(2,2))
  avg(x = "veh_age")
  avg(x = "veh_age_cat")
```


# Basic Data Manipulation
```{r }
summaries("area")
```

```{r }
dataCar$area_cat<-mapvalues(dataCar$area
                            , from = c("A","B","C","E")
                            , to = c("Others","Others","Others","Others"))
aggregate(dataCar$numclaims, by = list(dataCar[, "area_cat"]), FUN = mean)
summaries("area_cat")

par(mfrow=c(2,2))
  avg(x = "area")
  avg(x = "area_cat")
```

# Basic Data Manipulation
```{r }
summaries("agecat")
```

```{r }
dataCar$agecat_cat<-factor(ifelse(dataCar$agecat =='1','1'
                          ,ifelse(dataCar$agecat %in% c('5','6'),'5-6','Others')))
aggregate(dataCar$numclaims, by = list(dataCar[, "agecat_cat"]), FUN = mean)
summaries("agecat_cat")

par(mfrow=c(2,2))
  avg(x = "agecat")
  avg(x = "agecat_cat")
```

# 1. Poisson Regression
```{r }
formulas<-"numclaims ~ veh_body_cat"
poisson_reg1 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg1)
dataCar <- within(dataCar, veh_body_cat <- relevel(veh_body_cat, ref = 'Others'))
poisson_reg1 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg1)
```

# 1. Poisson Regression
```{r }
formulas<-"numclaims ~ veh_age_cat"
dataCar <- within(dataCar, veh_age_cat <- relevel(veh_age_cat, ref = 'Others'))
poisson_reg2 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg2)
```

# 1. Poisson Regression
```{r }
formulas<-"numclaims ~ area_cat"
dataCar <- within(dataCar, area_cat <- relevel(area_cat, ref = 'Others'))
poisson_reg3 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg3)
```

# 1. Poisson Regression
```{r }
formulas<-"numclaims ~ agecat_cat"
dataCar <- within(dataCar, agecat_cat <- relevel(agecat_cat, ref = 'Others'))
poisson_reg4 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg4)
```

# 1. Poisson Regression
```{r }
formulas<-"numclaims ~ veh_body_cat+veh_age_cat+area_cat+agecat_cat"
dataCar <- within(dataCar, agecat_cat <- relevel(agecat_cat, ref = 'Others'))
poisson_reg5 <- glm(formulas, data =dataCar, family=poisson)
summary(poisson_reg5)
```

# 2. Logistic Regression
```{r }
dataCar$numclaims_bin <- ifelse(dataCar$numclaims == 0, 0, 1)
table(dataCar$numclaims_bin)

formulas<-"numclaims_bin ~ veh_body_cat"
dataCar <- within(dataCar, veh_body_cat <- relevel(veh_body_cat, ref = 'Others'))
logistic_reg1 <- glm(formulas, data =dataCar, family=binomial)
summary(logistic_reg1)
```

# 2. Logistic Regression
```{r }
formulas<-"numclaims_bin ~ veh_age_cat"
dataCar <- within(dataCar, veh_age_cat <- relevel(veh_age_cat, ref = 'Others'))
logistic_reg2 <- glm(formulas, data =dataCar, family=binomial)
summary(logistic_reg2)
```

# 2. Logistic Regression
```{r }
formulas<-"numclaims_bin ~ area_cat"
dataCar <- within(dataCar, area_cat <- relevel(area_cat, ref = 'Others'))
logistic_reg3 <- glm(formulas, data =dataCar, family=binomial)
summary(logistic_reg3)
```

# 2. Logistic Regression
```{r }
formulas<-"numclaims_bin ~ agecat_cat"
dataCar <- within(dataCar, agecat_cat <- relevel(agecat_cat, ref = 'Others'))
logistic_reg4 <- glm(formulas, data =dataCar, family=binomial)
summary(logistic_reg4)
```

# 2. Logistic Regression
```{r }
formulas<-"numclaims_bin ~ veh_body_cat+veh_age_cat+area_cat+agecat_cat"
dataCar <- within(dataCar, agecat_cat <- relevel(agecat_cat, ref = 'Others'))
logistic_reg5 <- glm(formulas, data =dataCar, family=binomial)
summary(logistic_reg5)
```

# Model Selection
## Poisson Regression

* Nested Models

```{r }
anova(poisson_reg1,poisson_reg5, test="Chisq")
anova(poisson_reg2,poisson_reg5, test="Chisq")
anova(poisson_reg3,poisson_reg5, test="Chisq")
anova(poisson_reg4,poisson_reg5, test="Chisq")
```

* Non-Nested Models

```{r }
AICs<-c(poisson_reg1$aic,poisson_reg2$aic,poisson_reg3$aic,poisson_reg4$aic)
AICs
```

### Select 2 Poisson Regressions


# Model Selection
## Logistic Regression

* Nested Models

```{r }
anova(logistic_reg1,logistic_reg5, test="Chisq")
anova(logistic_reg2,logistic_reg5, test="Chisq")
anova(logistic_reg3,logistic_reg5, test="Chisq")
anova(logistic_reg4,logistic_reg5, test="Chisq")
```

* Non-Nested Models

```{r }
AICs<-c(logistic_reg1$aic,logistic_reg2$aic,logistic_reg3$aic,logistic_reg4$aic)
AICs
```

### Select 2 Logistic Regressions


# Model Selection
## Poisson Regression vs Logistic Regression

* You cannot use likelihood-based statistics like AIC to compare across models with different likelihood functions. 

* Difference in likelihood functions will account for the differences in the AIC probably more than differences in fit.
  1. Poisson regression: Poisson function
  2. Logistic regression: Bernoulli function. 
  
* Recommend broader approaches to choose the model
    + Predicted outcomes. MSE, cross-validation, etc.
    + Intuitive interpretation of coefficients

```{r }
MSEs<-c(mean(poisson_reg4$residuals^2),mean(poisson_reg5$residuals^2)
       ,mean(logistic_reg4$residuals^2),mean(logistic_reg5$residuals^2))
MSEs
```

### What model would you choose?

# Reference
https://en.wikipedia.org/wiki/Logistic_regression#Maximum_likelihood_estimation

https://en.wikipedia.org/wiki/Poisson_regression

https://cran.r-project.org/web/packages/insuranceData/insuranceData.pdf

http://stats.stackexchange.com/questions/139201/model-selection-can-i-compare-the-aic-from-models-of-count-data-between-linear

