---
title: 'Basic Data Manipulation'
author: "Brian A. Fannin"
output:
  slidy_presentation:
    css: My.css
    logo: images/cas_logo.png
    duration: 45
    fig_height: 3
    fig_width: 6
---

##

1. Brief review of vector operations
2. Data frames
3. Reading and writing data

## Vector review

```{r}
library(raw)
data(NJM_WC)

ay <- NJM_WC$AccidentYear
lag <- NJM_WC$Lag
dy <- NJM_WC$DevelopmentYear
ibnr <- NJM_WC$IBNR
paidLoss <- NJM_WC$CumulativePaid
incurredLoss <- NJM_WC$CumulativeIncurred

summary(paidLoss)
```

## Quick task

1. How much IBNR was held for development year 1997?
2. What was the case reserve held for accident year 1990 development lag 4?

## Answer

```{r}
sum(ibnr[dy == 1997])

whichCell <- ay == 1990 & lag == 4

incurredLoss[whichCell] - paidLoss[whichCell]

whichCell
```

## Data frames

A data frame is a set of vectors. This is basically the same concept as a table in Excel or a relational database.

Run the following lines:

```{r results='hide'}
summary(NJM_WC)
str(NJM_WC)
```

## Base operations - columns

Use `[ ]` and `$` to access elements of a data frame. Note that we need to include the comma to indicate that we're returning both rows and columns.

The following statements are equivalent:

```{r}
var1 <- NJM_WC$CumulativePaid
var2 <- NJM_WC[, "CumulativePaid"]
var3 <- NJM_WC[, 7]
```

## Base operations - rows

Filtering data by row works the same as with vectors. 

```{r}
upperTri <- NJM_WC$DevelopmentYear <= 1997
dfUpper <- NJM_WC[upperTri, ]
```

You can also use subset:

```{r}
dfUpper <- subset(NJM_WC, DevelopmentYear <= 1997)
```


## `dplyr`

| verb | outcome
|-------|-------------------------------------------|
| `select` | Returns columns that are named, or _excludes_ named columns|
| `filter` | Returns rows that meet a certain criteria |
| `mutate` | Creates a new column    |
| `group_by` | All subsequent operations are applied at the group level |
|  `summarise` | Aggregates a set of data to a single value |

The `%>%` operator will take the output of one operation as the input for the next operation.

## Quick example

```{r echo=FALSE}
suppressMessages(library(dplyr))
```

```{r }
library(dplyr)
dfEx <- NJM_WC %>% 
  group_by(AccidentYear) %>% 
  summarise(MaxPaid = max(CumulativePaid)
            , MinPaid = min(CumulativePaid)
            , AvgPaid = mean(CumulativePaid))
```

```{r results='asis', echo=FALSE}
pander::pandoc.table(dfEx, big.mark=",")
```

## `dplyr` in action

Identify 5 companies with the highest average paid for the first development year.

```{r}
data("wkcomp")

dfAvgPaid <- wkcomp %>%
  raw::CasColNames(FALSE) %>% 
  filter(Lag == 1) %>% 
  group_by(Company) %>% 
  summarise(AvgPaid = mean(CumulativePaid)) %>% 
  arrange(desc(AvgPaid)) %>% 
  slice(1:5)
```

```{r results='asis', echo=FALSE}
pander::pandoc.table(dfAvgPaid, big.mark=",")
```

## Task

1. Add a column to the New Jersey Manufacturer's data to indicate whether the cell is in the upper portion of the triangle.
2. Create a new data frame which includes only the upper triangle.

Additional tasks

1. Eliminate the GroupCode and Company columns from your data frame.
2. Form a time series which shows the ratio of net to direct earned premium. Note: you'll need to use the `unique` function.
3. Fit a linear model to this time series. Is there any relationship between cession ratio and accident year?

## Answer

```{r}
library(dplyr)

dfUpper <- NJM_WC %>% 
  mutate(Upper = DevelopmentYear <= 1997) %>% 
  filter(Upper)
```

## Additional answers

```{r}
dfMyData <- NJM_WC %>% 
  select(-GroupCode, -Company)

dfTimeSeries <- NJM_WC %>% 
  mutate(CessionPct = NetEP / DirectEP) %>% 
  unique()

myFit <- lm(CessionPct ~ AccidentYear, data = dfTimeSeries)
```

## Saving and reading data

Saving data

```{r}
write.csv(NJM_WC, "MyData.csv", row.names = FALSE)
```

Reading data

```{r}
dfMyData <- read.csv("MyData.csv", stringsAsFactors = FALSE)
```

## The working directory

Source of much confusion and frustration.

```{r eval=FALSE}
getwd()

setwd("SomeDirectory/SomeFolder")
```

## Task

1. Within your working directory create a new folder called "data".
2. Save the upper triangle data frame to this directory.

## Answer

```{r}
if (!dir.exists("./data/")) dir.create("./data/")

write.csv(dfUpper, "./data/upper.csv", row.names = FALSE)

dfUpper <- read.csv("./data/upper.csv", stringsAsFactors = FALSE)
```
