---
title: 'Basic Data Manipulation'
author: "Brian A. Fannin"
output:
  slidy_presentation:
    duration: 45
    fig_height: 3
    fig_width: 6
  beamer_presentation:
    colortheme: dolphin
---

##

1. Brief review of vector operations
2. Data frames
3. Reading and writing data

## Vector review

```{r}
library(raw)
data(NJM_WC)

ay <- NJM_WC$AccidentYear
lag <- NJM_WC$Lag
dy <- NJM_WC$DevelopmentYear
ibnr <- NJM_WC$IBNR
paidLoss <- NJM_WC$CumulativePaid
incurredLoss <- NJM_WC$CumulativeIncurred

summary(paidLoss)
```

## Quick task

1. How much IBNR was held for development year 1997?
2. What was the case reserve held for accident year 1990 development lag 4?

## Answer

```{r}
sum(ibnr[dy == 1997])

whichCell <- ay == 1990 & lag == 4

incurredLoss[whichCell] - paidLoss[whichCell]

whichCell
```

## Data frames

A data frame is a set of vectors. This is basically the same concept as a table in Excel or a relational database.

```{r results='hide'}
class(NJM_WC)

summary(NJM_WC)
str(NJM_WC)
```

## 

## Saving data

```{r}
write.csv(NJM_WC, "MyData.csv", row.names = FALSE)
```

## Reading data

```{r}
dfMyData <- read.csv("MyData.csv", stringsAsFactors = FALSE)

all.equal(NJM_WC, dfMyData)

class(NJM_WC)
```

## The working directory

## 

```{r}

```

## Task

* Subset the New Jersey Manufacturer's data to show only claims from 1990
* Form a data frame which does not show the GroupCode or Company columns
<!--
## The console

> - You can also execute lines from the console. 

> - You rarely want to do this.
-->