---
title: 'Visualization'
author: "Brian A. Fannin"
output:
  slidy_presentation:
    css: My.css
    duration: 45
    fig_height: 3
    fig_width: 6
  beamer_presentation:
    colortheme: dolphin
---

```{r echo=FALSE}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.pos="t"
)
```

## Base plotting

* Gets the job done with a minimum of fuss. 
* Advanced options and visual elements are possible, but a bit tough.

## A basic base plot

```{r}
dfUpper <- read.csv("./data/upper.csv", stringsAsFactors = FALSE)

plot(dfUpper$CumulativePaid, dfUpper$NetEP)
```

## ggplot2

ggplot2 developed by Hadley Wickham, based on the "grammar of graphics"
  
Particularly well suited for multi-dimensional, multivariate analysis.

Requires 3 things:

1. Data
2. Mapped aesthetics
3. Layers

## Data

```{r }
library(raw)
data(RegionExperience)
library(ggplot2)

basePlot <- ggplot(RegionExperience)
class(basePlot)
```

## Aesthetics

Aesthetics are anything visible on the plot. When an aesthetic is "mapped", it means that the data will define how the aesthetic appears. We'll map our aesthetics with the `aes` function.

```{r }
basePlot <- basePlot + aes(x = PolicyYear, y = NumClaims, color=Region)
```

## Adding layers

The `geom_*` functions add geometric shapes. 

```{r }
p <- basePlot + geom_line()
p
```

## Adding layers

```{r }
p <- basePlot + geom_point()
p
```

## Nothing wrong with adding two layers

```{r }
p <- basePlot + geom_point() + geom_line()
p
```

## One step

Typically we don't do this in steps.

```{r }
p <- ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims, group=Region, color=Region)) + geom_line()
p
```

## Layers have loads of parameters

```{r }
p <- basePlot + geom_bar(stat="identity", aes(fill = Region))
p
```

## Mmm, parameters

```{r }
p <- basePlot + geom_bar(stat="identity", position="dodge", aes(fill=Region))
p
```

## Layer parameters

How do I know which parameters do what? 

Help, google and stack overflow are your friends. Also, loads of books on the subject.

## Facets

```{r }
data(StateExperience)
p <- ggplot(StateExperience, aes(x = PolicyYear, y = NumClaims, color = State)) + geom_point() + facet_wrap(~ Region)
p <- p + theme(legend.position = "none")
p
```

## Statistical smoothers

```{r}
p <- ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims, group=Region, color=Region)) + geom_point()
p + geom_smooth(se = FALSE)
```

## Linear smoother

```{r }
p + geom_smooth(method = lm)
```

## Tasks

```{r}
data("wkcomp")
```

* Subset the wc reserving data to two companies
* Plot the cumulative paid against net earned premium. Split by development lag and company. You'll need to group and facet for this.
* Add a linear fit. 

Other tasks

* Draw a density plot which shows the lag 10 paid loss ratio separated by company.
* Are the responses normal? Is the variance the same?
* Plot the two companies in different facets

## Answer - Two companies

```{r results='hide', messages=FALSE}
suppressMessages(library(dplyr))
data("wkcomp")

dfTwo <- wkcomp %>% 
  raw::CasColNames(FALSE)

set.seed(1234)
dfTwo <- dfTwo %>% 
  filter(Company %in% sample(unique(dfTwo$Company), 2)) %>% 
  mutate(PaidLR = CumulativePaid / NetEP)
```

## Results by company, by lag

```{r}
plt <- ggplot(dfTwo, aes(NetEP, CumulativePaid, color = factor(Lag))) + geom_point() + facet_wrap(~Company, scales="free")
plt
```

## Results with a smoother

```{r }
plt + geom_smooth(method = lm, se=FALSE)
```

## Density

```{r}
pltDensity <- ggplot(filter(dfTwo, Lag == 10), aes(PaidLR, fill = Company)) + geom_density(alpha = 0.7)
pltDensity
```

## Density w/facet

```{r }
pltDensity + facet_wrap(~ Company)
```

<!--
## Maps!

```{r }
library(maps)
map('state')
```

## Hurricanes

```{r }
data(Hurricane)
 
dfKatrina = subset(Hurricane, Name == 'KATRINA')
dfKatrina = dfKatrina[dfKatrina$Year == max(dfKatrina$Year), ]

 
dfHugo = subset(Hurricane, Name == 'HUGO')
dfHugo = dfHugo[dfHugo$Year == max(dfHugo$Year), ]
 
dfDonna = Hurricane[Hurricane$Name == 'DONNA', ]
dfDonna = dfDonna[dfDonna$Year == max(dfDonna$Year), ]
```

## Hurricanes

```{r }
map('state')
points(dfKatrina$Longitude, dfKatrina$Latitude, pch=19, col = 'red')
points(dfHugo$Longitude, dfHugo$Latitude, pch = 19, col = 'blue')
points(dfDonna$Longitude, dfDonna$Latitude, pch = 19, col = 'green')
```

## Choropleths -->