---
title: 'Visualization'
author: "Brian A. Fannin"
output:
  slidy_presentation:
    css: My.css
    duration: 45
    fig_height: 3
    fig_width: 6
  beamer_presentation:
    colortheme: dolphin
---

## 

* base plotting
* ggplot2

## Base plotting

* Gets the job done with a minimum of fuss. 
* Advanced options and visual elements are possible, but a bit tough.

##

```{r}

```


## ggplot2

ggplot2 developed by Hadley Wickham, based on the "grammar of graphics"
  
Particularly well suited for multi-dimensional, multivariate analysis.

Requires 3 things:

1. Data
2. Aesthetics
3. Layers

## Data

```{r }
library(raw)
data(RegionExperience)
library(ggplot2)

basePlot <- ggplot(RegionExperience)
class(basePlot)
```

## Aesthetics

Aesthetics are anything visible on the plot. Added with the `aes` function. Data will be mapped to visual elements.

```{r }
basePlot <- basePlot + aes(x = PolicyYear, y = NumClaims, group=Region, color=Region, fill=Region)
```

## Adding layers

The `geom_*` functions add geometric shapes. 

```{r }
p <- basePlot + geom_line()
p
```

## Adding layers

```{r }
p <- basePlot + geom_point()
p
```

## Nothing wrong with adding two layers

```{r }
p <- basePlot + geom_point() + geom_line()
p
```

## Adding layers

```{r }
p <- basePlot + geom_bar(stat="identity")
p
```

Here `stat` is a statistical transformation.

## One step

Typically we don't do this in steps.

```{r }
p <- ggplot(RegionExperience, aes(x = PolicyYear, y = NumClaims, group=Region, color=Region)) + geom_line()
p
```

## Facets
```{r }
p <- basePlot + geom_line() + facet_wrap(~ Region)
p
```

## Facets

That plot was a bit boring.

```{r }
data(StateExperience)
p = ggplot(StateExperience, aes(x = PolicyYear, y = NumClaims, color = State)) + geom_point() + facet_wrap(~ Region)
p
```

```{r eval=FALSE}
p <- ggplot(subset(StateExperience, Region == "South"), aes(x = PolicyYear, y = NumClaims, label=Postal))
p <- p + geom_point() + facet_wrap(~ Postal)
p
```

## Tasks

```{r}
data("wkcomp")
```

* Subset the wc reserving data to two companies
* Plot the cumulative paid against net earned premium. Ensure that each company is distinguishable.
* Add a linear fit. 

## Answer

```{r}

```


## Other tasks

* Draw a density plot which shows the lag 1 paid loss ratio separated by company.
* Are the responses normal? Is the variance the same?
* Plot the two companies in different facets and 

## Other answers

```{r}
library(dplyr)
data("wkcomp")

dfTwo <- wkcomp %>% 
  raw::CasColNames(FALSE)

set.seed(1234)
dfTwo <- dfTwo %>% 
  filter(Company %in% sample(unique(dfTwo$Company), 2)
         , Lag == 1) %>% 
  mutate(PaidLR = CumulativePaid / NetEP)

pltDensityOne <- ggplot(dfTwo, aes(PaidLR, fill = Company)) + geom_density(alpha = 0.7)
pltDensityOne
```


<!--
## Maps!

```{r }
library(maps)
map('state')
```

## Hurricanes

```{r }
data(Hurricane)
 
dfKatrina = subset(Hurricane, Name == 'KATRINA')
dfKatrina = dfKatrina[dfKatrina$Year == max(dfKatrina$Year), ]

 
dfHugo = subset(Hurricane, Name == 'HUGO')
dfHugo = dfHugo[dfHugo$Year == max(dfHugo$Year), ]
 
dfDonna = Hurricane[Hurricane$Name == 'DONNA', ]
dfDonna = dfDonna[dfDonna$Year == max(dfDonna$Year), ]
```

## Hurricanes

```{r }
map('state')
points(dfKatrina$Longitude, dfKatrina$Latitude, pch=19, col = 'red')
points(dfHugo$Longitude, dfHugo$Latitude, pch = 19, col = 'blue')
points(dfDonna$Longitude, dfDonna$Latitude, pch = 19, col = 'green')
```

## Choropleths -->