---
title: 'ChainLadder'
author: "Rajesh Sahasrabuddhe"
output:
 
  ioslides_presentation:
    duration: 45
    fig_height: 3
    fig_width: 6
    transition: faster
    widescreen: yes
  beamer_presentation:
    colortheme: dolphin
---

## What is this 'ChainLadder'

The `ChainLadder` package provides various statistical methods which are typically used for the estimation of outstanding claims reserves in general insurance, including those to estimate the claims development results as required under Solvency II.  

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='asis'}
library(ChainLadder)
citation('ChainLadder')$textVersion
```

## What is this 'ChainLadder'

Authors:  
<div class="columns-2">
- [Markus Gesmann](http://www.magesblog.com/)  
- [Dan Murphy, FCAS, MAAA](http://trinostics.blogspot.com/)  
- [Yanwei (Wayne) Zhang, FCAS](https://www.linkedin.com/in/yanwei-wayne-zhang-314597ab)  
- [Alessandro Carrato, FIA](https://www.linkedin.com/in/alessandrocarrato)  
- [Fabio Concina](https://www.linkedin.com/in/fabioconcina)  
- [Prof. Dr. Mario V. Wüthrich](https://people.math.ethz.ch/~wueth/)  
</div>


Github site:  
- [https://github.com/mages/ChainLadder](https://github.com/mages/ChainLadder)


## The Triangle Class

Functions to ease the work with triangle shaped matrix data. A 'triangle' is a matrix with some generic functions. Triangles are usually stored in a 'long' format in data bases. The function `as.triangle` can transform a data.frame into a triangle shape.

```{r}
head(GenInsLong)
```

## The Triangle Class

```{r}
class(GenInsLong)
gen_ins_tri <- as.triangle(Triangle = GenInsLong, origin = "accyear", dev = "devyear", 
  value = "incurred claims")
class(gen_ins_tri)

```

## The Triangle Class

```{r}
str(gen_ins_tri)
```

## The Triangle Class
```{r}
gen_ins_tri[,1:8]
```

## Excercise 1
1. Load the `ChainLadder` and `raw` packages
2. For the `NJM_WC` dataset:
    - Create an `R` object with class `triangle` and `matrix` with the incurred loss data
    - Remove all records with `DevelopmentYear` greater than 1997  
    Extra Credit
    - Create an `R` object with class `data.frame` or `tbl_df` with the observed development factors for further examination
    - Using plotting (i.e. data visualization) tools only identify the accident year that results in the largest 2:3 development factor.
    
## Excercise 1

```{r, message=FALSE, warning=FALSE}
library(ChainLadder)
library(raw)
library(magrittr)
library(tibble)
library(tidyr)
library(dplyr)
library(ggplot2)

class(NJM_WC)

njm_wc_tri <- NJM_WC %>% 
  dplyr::filter(DevelopmentYear <= 1997) 
class(njm_wc_tri)

```
## Excercise 1
```{r}
njm_wc_tri <- as.triangle(Triangle = njm_wc_tri, origin = 'AccidentYear', 
  dev = 'Lag', value =  'CumulativeIncurred'); njm_wc_tri[,1:8]

```

    
## Excercise 1    
```{r}
class(njm_wc_tri)
dim(njm_wc_tri)
njm_wc_df <- njm_wc_tri[,2:10] / njm_wc_tri[,1:9] %>% 
  as_tibble()
njm_wc_df <- dplyr::mutate(njm_wc_df, accident_year = as.character(1988:1997))
names(njm_wc_df)[1:9] <- paste(1:9, "to", 2:10)

njm_wc_df <- gather(njm_wc_df, key = interval, value = development_factor, 1:9, 
  na.rm = TRUE)

head(njm_wc_df)

ggplot(data = njm_wc_df, mapping = aes(x = interval, y = development_factor, 
  color = accident_year)) + geom_point()

```


