---
title: 'ChainLadder'
author: "Rajesh Sahasrabuddhe"
output:
  ioslides_presentation:
    logo: images/cas_logo.png
    duration: 45
    fig_height: 3
    fig_width: 6
    transition: faster
    widescreen: yes
  beamer_presentation:
    colortheme: dolphin
---

```{r}
library(methods)
```

## Agenda

- The ChainLadder package
- The Triangle Class
- Models available in ChainLadder
- Other Topics


## What is this 'ChainLadder'?

The `ChainLadder` package provides various statistical methods which are typically used for the estimation of outstanding claims reserves in general insurance, including those to estimate the claims development results as required under Solvency II.  

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='asis'}
library(ChainLadder)
citation('ChainLadder')$textVersion
```

## What is this 'ChainLadder'?

Authors:  
<div class="columns-2">
- [Markus Gesmann](http://www.magesblog.com/)  
- [Dan Murphy, FCAS, MAAA](http://trinostics.blogspot.com/)  
- [Yanwei (Wayne) Zhang, FCAS](https://www.linkedin.com/in/yanwei-wayne-zhang-314597ab)  
- [Alessandro Carrato, FIA](https://www.linkedin.com/in/alessandrocarrato)  
- [Fabio Concina](https://www.linkedin.com/in/fabioconcina)  
- [Prof. Dr. Mario V. Wüthrich](https://people.math.ethz.ch/~wueth/)  
</div>

Github site:

- [https://github.com/mages/ChainLadder](https://github.com/mages/ChainLadder)


## The Triangle Class

Functions to ease the work with triangle shaped matrix data. A 'triangle' is a matrix with some generic functions. Triangles are usually stored in a 'long' format in data bases. The function `as.triangle` can transform a data.frame into a triangle shape.

```{r}
head(GenInsLong)
```

## The Triangle Class

```{r}
class(GenInsLong)
gen_ins_tri <- as.triangle(Triangle = GenInsLong, origin = "accyear", dev = "devyear", 
  value = "incurred claims")
class(gen_ins_tri)

```

## The Triangle Class

```{r}
str(gen_ins_tri)
```

## The Triangle Class
```{r}
gen_ins_tri[,1:8]
```
 
## Exercise 1
1. Load the `ChainLadder` and `raw` packages
2. For the `NJM_WC` dataset:
    - Create an `R` object with class `triangle` and `matrix` with the incurred loss data
    - Remove all records with `DevelopmentYear` greater than 1997  
<div class="red" style: color="red"><br>Extra Credit</div>
    - Create an `R` object with class `data.frame` or `tbl_df` with the observed development factors for further examination
    - Using plotting (i.e. data visualization) tools only, identify the accident year that results in the largest 2:3 development factor.
    
## Exercise 1

```{r, message=FALSE, warning=FALSE}
library(ChainLadder)
library(raw)
library(magrittr)
library(tibble)
library(tidyr)
library(dplyr)
library(ggplot2)

class(NJM_WC)

```

## Exercise 1
```{r}
njm_wc_tri <- NJM_WC %>% 
  dplyr::filter(DevelopmentYear <= 1997) 
class(njm_wc_tri)

```



## Exercise 1
```{r}
njm_wc_tri <- as.triangle(Triangle = njm_wc_tri, origin = 'AccidentYear', 
  dev = 'Lag', value =  'CumulativeIncurred'); njm_wc_tri[,1:8]

```

## Exercise 1    
```{r}
class(njm_wc_tri)
dim(njm_wc_tri)
njm_wc_df <- njm_wc_tri[,2:10] / njm_wc_tri[,1:9] %>% 
  as_tibble()
njm_wc_df <- dplyr::mutate(njm_wc_df, accident_year = as.character(1988:1997))
names(njm_wc_df)[1:9] <- paste(1:9, "to", 2:10)
```

## Exercise 1    
```{r}
njm_wc_df <- gather(njm_wc_df, key = interval, value = development_factor, 1:9, 
  na.rm = TRUE)

head(njm_wc_df)

```

## Exercise 1    
```{r}
ggplot(data = njm_wc_df, mapping = aes(x = interval, y = development_factor, 
  color = accident_year)) + geom_point()

```


## Actuarial Projection Models

Models Available in `ChainLadder`:
<div class="columns-2">
- `BootChainLadder` (`CDR`)
- `ClarkCapeCod` 
- `ClarkLDF` 
- `glmReserve`
- `MackChainLadder` (`CDR`)  
- `tweedieReserve`  
</div>
- `MultiChainLadder` (Join2Fits)
- `MunichChainLadder` (Requires Paid and Incurred)
- `PaidIncurredChain` (Requires Paid and Incurred)

## How do I use these models?

1. Read the help function  
    - Understand the arguments  
2. Apply the function and assign the result to an object  
3. Understand the output  
    - `str`
    - `summary`
    - `plot`  
4. Extract what you need 
    - S3 class `[`
    - S4 class `$`

## BootChainLadder

`BootChainLadder(Triangle, R = 999, process.distr=c("gamma", "od.pois"))`

- The BootChainLadder procedure provides a predictive distribution of reserves or IBNRs for a cumulative claims development triangle.

- `process.distr` - character string indicating which process distribution to be assumed. One of "gamma" (default), or "od.pois" (over-dispersed Poisson), can be abbreviated

- _England, PD and Verrall, RJ. Stochastic Claims Reserving in General Insurance (with discussion)_, British Actuarial Journal _8, III. 2002_

## Excercise 2
Run the following code to create a paid triangle for other liability:
```{r}
liab_tri <- as.triangle(Triangle = 
    othliab[othliab$GRNAME == 'State Farm Mut Grp' & 
        othliab$DevelopmentYear <= 1997, ],
  origin = 'AccidentYear', dev = 'DevelopmentLag', value = 'CumPaidLoss_h1')
```

## Excercise 2
For the ClarkLDF method estimate with a tail that is:  

1. 20 years in length
2. measured from the beginning of the accident year
3. modeled using a loglogistic growth curve

## Excercise 2
Provide the following:

1. Total 'IBNR' (Unpaid)
2. The indicated cumulative LDF for 1989 for the specified growth function (without regards to the selected length of the tail)
3. The LDF truncated to 20 years for 1989  
<div class="red" style: color="red"><br>Extra Credit</div> 
4. The indicated cumulative development factor at age 15 (with and without truncation)

## Excercise 2
```{r}
clark <- ClarkLDF(Triangle = liab_tri, maxage = 20, adol = FALSE)
clark$Total$FutureValue
clark$Ldf[2]
clark$TruncatedLdf[2]
```
## Excercise 2
```{r}
unname(1 + (clark$THETAG[2]/15)^clark$THETAG[1])
unname(1 + (clark$THETAG[2]/15)^clark$THETAG[1] / 
    1 + (clark$THETAG[2]/20)^clark$THETAG[1])
```


## Tails
Both the `ClarkCapeCod` and `ClarkLDF` fit growth functions to extrapolate the development pattern. 

`tail ~ c("G", "maxage")`

See the Clark paper for the forms of the extrapolation model.

## Selected Loss Development Factors
The `CLFMdelta` function
```{r}
init_df <- MackChainLadder(Triangle = liab_tri)$f
sel_df <- init_df
sel_df[3] <- 1.5
alphas <- CLFMdelta(Triangle = liab_tri, selected = sel_df[1:9])
full_tri <- predict(chainladder(liab_tri, delta = alphas))
```
## Selected Loss Development Factors

```{r}
full_tri[7:10,1:5]
full_tri[,4] / full_tri[,3]
```

## Selected Loss Development Factors
This has the potential to support the variability calculation from the [CLFM paper](http://www.variancejournal.org/issues/06-02/143.pdf)

[Github issue](https://github.com/mages/ChainLadder/issues/23)


## Quarterly Triangles
The `qpaid` and `qincurred` are included in ChainLadder to demonstrate how to deal with triangles with quarterly evaluations and annual periods.
```{r}
# ?qpaid #not run
dim(qpaid)
## MackChainLadder expects a quadratic matrix so let's expand 
## the triangle to a quarterly origin period.
n <- ncol(qpaid) # Number of quarterly valuations
Paid <- matrix(NA, n, n) # create a matrix with NAs
Paid[seq(from = 1, to = n, by = 4),] <- qpaid # fill in every 4th row with the annual data
```

## Questions??

Rajesh Sahasrabuddhe
215 246 1028
[Write me!](mailto:rajesh.sahasrabuddhe@oliverwyman.com)
Final