---
title: 'Reserving w/o ChainLadder'
author: "Brian A. Fannin"
output:
  slidy_presentation:
    css: My.css
    duration: 45
    fig_height: 3
    fig_width: 6
  beamer_presentation:
    colortheme: dolphin
---

```{r echo = FALSE}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.pos="t"
)
```

## Reserving w/o ChainLadder

1. What is the additive model?
2. Form incremental and prior cumulative amounts
3. Fit a linear model
4. Compare our fits

## The additive model

* Incremental development amounts are a linear function of some exposure measure.
* Similar to Bornhuetter-Ferguson, but doesn't require an a priori loss ratio assumption.
* Don't expect it to work as well if earned premium has not been on-leveled.

## Form incrementals

```{r echo=FALSE, results='hide'}
suppressPackageStartupMessages(library(dplyr))
```


```{r}
library(dplyr)

dfUpper <- read.csv("./data/upper.csv", stringsAsFactors = FALSE) %>% 
  arrange(AccidentYear, Lag) %>% 
  mutate(IncrementalPaid = CumulativePaid - lag(CumulativePaid)
         , IncrementalPaid = ifelse(Lag == 1, CumulativePaid, IncrementalPaid)
         , PriorCumulativePaid = lag(CumulativePaid)
         , PriorCumulativePaid = ifelse(Lag == 1, NA, PriorCumulativePaid))
```

## Check our data

```{r }
dfCheck <- dfUpper %>% 
  group_by(AccidentYear) %>% 
  mutate(Check = cumsum(IncrementalPaid)) %>% 
  filter(Check != CumulativePaid)
```

## Lag is a category - not an integer

```{r}
dfUpper <- dfUpper %>% 
  mutate(LagGroup = factor(Lag))
```


## Observe our data - additive

```{r echo=FALSE}
suppressPackageStartupMessages(library(ggplot2))
```


```{r}
library(ggplot2)

plt <- ggplot(dfUpper, aes(NetEP, IncrementalPaid, color = LagGroup)) + geom_point()
plt <- plt + geom_smooth(method = "lm", se = FALSE)
plt
```

## Your turn

Create a similar plot that uses `PriorCumulativePaid` as the predictor variable.

## Observe our data - chain ladder

```{r}

plt <- ggplot(dfUpper, aes(PriorCumulativePaid, IncrementalPaid, color = LagGroup)) + geom_point()
plt <- plt + geom_smooth(method = "lm", se = FALSE)
plt
```

## Fit a model

```{r}
fitPaidAM <- lm(data = dfUpper
                , IncrementalPaid ~ 0 + NetEP:LagGroup)
```

## 

```{r}
summary(fitPaidAM)
```

## Your turn

* Construct a linear fit which uses prior cumulative paid as the predictor

## 

```{r}
fitPaidCL <- lm(data = dfUpper
                , IncrementalPaid ~ 0 + PriorCumulativePaid:LagGroup)
```

##

```{r}
summary(fitPaidCL)
```

## Predictions and residuals

```{r}
dfUpper <- dfUpper %>% 
  mutate(PredictedPaidAM = predict(fitPaidAM)
         , ResidualPaidAM = residuals(fitPaidAM)
         , R_StandardPaidAM = rstandard(fitPaidAM))
```

## Plots

* Plot the residuals as a function of the predicted value.
* Plot the residuals as a function of accident year, lag and development year

## Plots (cont'd)

```{r}
pltResid <- ggplot(dfUpper, aes(PredictedPaidAM, R_StandardPaidAM)) + geom_point()
pltResid

pltResid + geom_smooth(method = "lm")
```


## Model comparison

* Which model has a better r^2?
* Which model has a better F statistic?
* What should we consider when comparing our models? How could adjust for this?


## Model comparison (cont'd)

We can inspect the printed output from `summary`. You can also access diagnostic tests directly.

```{r}
summary(fitPaidAM)$fstatistic
summary(fitPaidCL)$fstatistic
```

## Tasks

* Form incrementals and priors for the completed triangle for NJM_WC
* Predict values for each cell using the fit you constructed _on the upper triangle_.
* Which model performs better?

## Answers

```{r}
library(raw)
data(NJM_WC)
```

```{r }
NJM_WC <- NJM_WC %>% 
  arrange(AccidentYear, Lag) %>% 
  mutate(IncrementalPaid = CumulativePaid - lag(CumulativePaid), 
         IncrementalPaid = ifelse(Lag == 1, CumulativePaid, IncrementalPaid),
         PriorCumulativePaid = dplyr::lag(CumulativePaid), 
         PriorCumulativePaid = ifelse(Lag == 1, NA, PriorCumulativePaid),
         LagGroup = factor(Lag)) 
```

## Answers - new predictions

```{r }
NJM_WC$PredictAM <- predict(fitPaidAM, newdata = NJM_WC)
NJM_WC$PredictCL <- NA
LagGt1 <- NJM_WC$Lag > 1
NJM_WC$PredictCL[LagGt1] <- predict(fitPaidCL, newdata = NJM_WC[LagGt1, ])
```

## Answers - model comparisons

```{r}
NJM_WC <- NJM_WC %>% 
  mutate(AM_Error = PredictAM - IncrementalPaid, 
         CL_Error = PredictCL - IncrementalPaid, 
         Upper = DevelopmentYear <= 1997)

sqrt(sum(NJM_WC$AM_Error[!NJM_WC$Upper]^2))
sqrt(sum(NJM_WC$CL_Error[!NJM_WC$Upper]^2))
```

## MRMR

The functionality above is incorporated in the MRMR package. The beta version is on GitHub and may be installed using the following:

```{r eval=FALSE}
devtools::install_github("PirateGrunt/MRMR")
```

